DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-04-18 paper/comparison.tex
revision 4.E (2nd)
comparsion and conclusion revision

@@

Figure~\ref{fig:sys_core} shows the average control loop completion
timing of the four system configurations we tested as a function of
the number of CPU cores used.
First, both the Intel UP and Jetson TX2 exhibit superior performance when
compared with the Raspberry Pi 3. 
When all four CPU cores are used, the Intel UP is 1.33X faster than
the Pi 3, while the TX2 (CPU) and TX2 (GPU) are 2.79X and 4.16X times faster,
respectively, than the Pi 3. 
As a result, they are all able to satisfy the 33.$\overline{\mbox{33}}$ ms 
WCET by a clear margin,
%%  (in the worst case, the UP Board would 
%% still meet it by $\sim$20 ms) --> where's the data showing this?
and, in the case of the TX2, 50 Hz or even 100 Hz real-time control is
feasible with the help of its GPU. Another observation is that the TX2
(GPU) performance does not change much, as most of the neural network 
computation is done by the GPU.

@@

The Intel UP board and Jetson TX2 also perform much better when multiple DNN models
are co-scheduled. Figure~\ref{fig:sys_model} shows the results of the
multi-model co-scheduling experiment. Once again, they can comfortably
satisfy 30 Hz real-time performance for all of the co-scheduled DNN control
loops, and in the case of the TX2 (GPU), 100 Hz real-time control is still
feasible. Given that the GPU must be shared among the co-scheduled DNN
models, the results suggest that the TX2's GPU has sufficient capacity to
accomodate multiple instances of the DNN models we tested.

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/compare_benchmark}
  \caption{Average control loop execution time in the presence of an
    increasing number of memory intensive applications on idle CPU cores.}
  \label{fig:sys_bench}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=.45\textwidth]{figs/compare_l2missrate}
  \caption{L2 miss rates in the presence of an increasing number of 
			memory intensive applications on idle CPU cores.}
  \label{fig:sys_l2miss}
\end{figure}

Finally, we compare the effect of co-scheduling memory bandwidth
intensive synthetic benchmarks on the DNN control loop timing. 
Figure~\ref{fig:sys_bench} shows the results. As discussed in 
Section~\ref{sec:eval-memhog}, we observed dramatic execution time
increases, up to 11.6X, in Raspberry Pi 3 as we increased the number of
co-scheduled tasks. We also observe increased control loop execution
timing in the Intel Up and Jetson TX2, but the degree of the increase is 
not as dramatic as the Pi 3. Compared to their respective solo timings 
(i.e., the model runs on a single core in isolation), Intel UP suffers up to
2.3X execution time increase; TX2 (CPU) and TX2 (GPU) suffer up to
2.2X and 2.1X increases, respectively. This is somewhat surprising
because the Raspberry Pi 3's cores are in-order architecture based while
the cores in the Intel Up and NVIDIA TX2 are out-of-order architecture
based, and that the memory intensive tasks on out-of-order cores can
generate more memory traffic. We believe that this is because the
memory subsystems in the Intel UP and TX2 platforms provide higher
performance than the memory subsystem of the Pi 3 as suggested from
the measured memory bandwidth results in Table~\ref{tbl:platforms}
('Measured Bandwidth').

Another interesting observation is that the TX2 (GPU) also suffers
considerable execution time increase (2.1X) despite the fact that the
co-scheduled synthetic tasks do not utilize the GPU. In other words,
the DNN model has dedicated access to the GPU. This is, however, a
known characteristic of integrated CPU-GPU architecture based
platforms where both CPU and GPU share the same memory
subsystem~\cite{Ali2017}. As a result, the TX2 (GPU) fails to meet the
10ms deadline for 100 Hz control that would have been feasible if
there was no contention between the CPU cores and the GPU.

Figure~\ref{fig:sys_l2miss} shows the L2 miss rates of DNN control
loop. The key takeaway of this result is that the L2 miss rate is a
poor indicator to the execution time of the control loop. As shown
already, the Raspberry Pi 3's increased L2 misses do not fully explain
increases in processing times. In case of the Intel UP, as the number of
interfering Bandwidth benchmark instances increases, the L2 miss rates
drop slightly. In the case of the TX2 (CPU), from solo to one co-runner,
L2 miss rates almost double, but then the execution time
increase is a little less than 10\%. However, when there are
co-runners, even though the L2 miss rates are largely unchanged, the
DNN processing time increases significantly. The TX2 (GPU) also shows
similar behavior. All of these results essentially point to the fact
that DNN inferencing workload is largely memory performance sensitive
but not L2 cache space sensitive.

%% However, the L2 miss rates of the Pi 3 and TX2 platforms follow the same expected 
%% pattern, as can be seen in Figure~\ref{fig:sys_l2miss}. Namely, the 
%% percentage of L2 miss rates increases are more 
%% memory intensive co-runners are introduced. This was not the case for 
%% the UP board, where the model running on two cores generated the highest
%% percentage of L2 cache misses. This is most likely due to the CPU 
%% architecture of the UP's CPU which is Intel based, whereas both the Pi 3
%% and TX2 have ARM based CPUs.

>>>>

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.45\textwidth]{figs/a53_vs_a7}
%%   \caption{DNN model performance on different in-order
%% ARM Cortex cores. \fixme{this may be better in the plaform comparison section}.}
%%   \label{fig:a53_vs_a7}
%% \end{figure}

@@

Figure~\ref{fig:sys_core} shows the average control loop completion
timing of the four system configurations we tested as a function of
the number of CPU cores used.
First, both the Intel UP and Jetson TX2 exhibit superior performance when
compared with the Raspberry Pi 3. 
When all four CPU cores are used, the Intel UP is 1.33X faster than
the Pi 3, while the TX2 (CPU) and TX2 (GPU) are 2.79X and 4.16X times faster,
respectively, than the Pi 3. 
As a result, they are all able to satisfy the 33.3 ms 
WCET by a clear margin,
%%  (in the worst case, the UP Board would 
%% still meet it by $\sim$20 ms) --> where's the data showing this?
and, in the case of the TX2, 50 Hz or even 100 Hz real-time control is
feasible with the help of its GPU. Another observation is that the TX2
(GPU) performance does not change much, as most of the neural network 
computation is done by the GPU.

@@

Figure~\ref{fig:sys_model} shows the results of the
multi-model co-scheduling experiment. Once again, they can comfortably
satisfy 30 Hz real-time performance for all of the co-scheduled DNN control
loops, and in the case of the TX2 (GPU), 100 Hz real-time control is still
feasible. Given that the GPU must be shared among the co-scheduled DNN
models, the results suggest that the TX2's GPU has sufficient capacity to
accomodate multiple instances of the DNN models we tested.

\begin{figure}[h]
  \centering
  \begin{subfigure}{0.4\textwidth}
    \includegraphics[width=\textwidth]{figs/compare_benchmark_read}
    \caption{BwRead}
    \label{fig:sys_bench_read}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.4\textwidth}
    \includegraphics[width=\textwidth]{figs/compare_benchmark}
    \caption{BwWrite}
    \label{fig:sys_bench_write}    
  \end{subfigure}
  \caption{Average processing time vs. the number of memory
    memory intensive co-runners.}
  \label{fig:sys_bench}    
\end{figure}

Figure~\ref{fig:sys_bench} shows the results of synthetic memory
intensive task co-scheduling experiment. 
For read co-runners (BwRead), the performance of all platforms
gradually decreased as additional BwRead instances are introduced: up
to 1.6X for Pi 3, up to 1.3X for Intel UP, and up to 1.6X and 2.2X for the TX2
(CPU) and TX2(GPU), respectively.
For write co-runners (BwWrite), however, we observe generally more
noticieable execution time increases. In particular, as we discussed
earlier in Section~\ref{sec:evaluation}, the Pi 3 suffers up to 11.6X
execution time increase. We also observe increased control loop
execution timing in the Intel Up and Jetson TX2, although the degree
of the increase is not as dramatic as the Pi 3.

%% Compared to their respective solo timings  
%% (i.e., the model runs on a single core in isolation), Intel UP suffers up to
%% 2.3X execution time increase; TX2 (CPU) and TX2 (GPU) both suffer up to
%% 2.3X increases. This is somewhat surprising
%% because the Raspberry Pi 3's cores are in-order architecture based while
%% the cores in the Intel Up and NVIDIA TX2 are out-of-order architecture
%% based, and that the memory intensive tasks on out-of-order cores can
%% generate more memory traffic.
%% We believe that this is because the
%% memory subsystems in the Intel UP and TX2 platforms provide higher
%% performance than the memory subsystem of the Pi 3 as suggested from
%% the measured memory bandwidth results in Table~\ref{tbl:platforms}
%% ('Measured Bandwidth').

Another interesting observation is that the TX2 (GPU) also suffers
considerable execution time increase (2.3X) despite the fact that the
co-scheduled synthetic tasks do not utilize the GPU. In other words,
the DNN model has dedicated access to the GPU. This is, however, a
known characteristic of integrated CPU-GPU architecture based
platforms where both CPU and GPU share the same memory
subsystem~\cite{Ali2017}. As a result, the TX2 (GPU) fails to meet the
10ms deadline for 100 Hz control that would have been feasible if
there was no contention between the CPU cores and the GPU.

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.45\textwidth]{figs/compare_readl2}
%%   \caption{L2 miss rates in the presence of an increasing number of 
%% 			memory intensive read applications on idle CPU cores.}
%%   \label{fig:sys_readl2miss}
%% \end{figure}

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.45\textwidth]{figs/compare_l2missrate}
%%   \caption{L2 miss rates in the presence of an increasing number of 
%% 			memory intensive write applications on idle CPU cores.}
%%   \label{fig:sys_l2miss}
%% \end{figure}

%% Figures~\ref{fig:sys_readl2miss} and ~\ref{fig:sys_l2miss} show
%% the L2 miss rates of DNN control loops when read and write co-runners
%% are introduced, respectively. The key takeaway of this result is that the 
%% L2 miss rate is a
%% poor indicator to the execution time of the control loop. As shown
%% already, the Raspberry Pi 3's increased L2 misses do not fully explain
%% increases in processing times. In case of the Intel UP, as the number of
%% interfering Bandwidth benchmark instances increases, the L2 miss rates
%% drop slightly. In the case of the TX2 (CPU), from solo to one co-runner,
%% L2 miss rates almost double, but then the execution time
%% increase is a little less than 10\%. However, when there are
%% co-runners, even though the L2 miss rates are largely unchanged, the
%% CNN processing time increases significantly. The TX2 (GPU) also shows
%% similar behavior. All of these results essentially point to the fact
%% that DNN inferencing workload is largely memory performance sensitive
%% but not L2 cache space sensitive.

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.45\textwidth]{figs/compare_readbus}
%%   \caption{Bus accesses in the presence of an increasing number of 
%% 			memory intensive read applications on idle CPU cores.}
%%   \label{fig:sys_readbus}
%% \end{figure}

%% \begin{figure}[h]
%%   \centering
%%   \includegraphics[width=.45\textwidth]{figs/compare_writebus}
%%   \caption{Bus accesses in the presence of an increasing number of 
%% 			memory intensive write applications on idle CPU cores.}
%%   \label{fig:sys_writebus}
%% \end{figure}

%% \color{red}FIXME: I believe the bus access results for the Intel UP are wrong. 
%% I couldn't get the IDI off-core response event to work, so I used the bus-cycles 
%% perf event, which I think is incorrect.

%% \color{black}Figures~\ref{fig:sys_readbus} and ~\ref{fig:sys_writebus} show the
%% normalized bus accesses for each of the platforms as read and write
%% co-runners are introduced, respectively. Unlike the L2 miss rates of the
%% model, the total number of bus accesses more closely align with the 
%% execution time increases that are seen in each of the platforms. For the Pi 3, both 
%% the execution time and bus accesses increase at exponential rates. In the TX2(GPU), 
%% the bus accesses increase in a similar linear fashion to the execution time. The main outlier
%% is the TX2(CPU) which has bus accesses increase at a faster rate similar to that of the Pi 3.
%% However, the bus accesses seem to increase linearly, which is also seen in the
%% execution time. As a result, we find that the total number of bus accesses have a higher 
%% correlation to the model execution time than L2 miss rates.

%% %% However, the L2 miss rates of the Pi 3 and TX2 platforms follow the same expected 
%% %% pattern, as can be seen in Figure~\ref{fig:sys_l2miss}. Namely, the 
%% %% percentage of L2 miss rates increases are more 
%% %% memory intensive co-runners are introduced. This was not the case for 
%% %% the UP board, where the model running on two cores generated the highest
%% %% percentage of L2 cache misses. This is most likely due to the CPU 
%% %% architecture of the UP's CPU which is Intel based, whereas both the Pi 3
%% %% and TX2 have ARM based CPUs.

### 2018-04-18 paper/conclusion.tex
comparsion and conclusion revision

As future work, 
%% we plan to apply shared DRAM resource management
%% techniques~\cite{Yun2013,yun2014rtas} on the DeepPicar platform and
%%evaluate their impact on the real-time performance of the system. 
we plan to improve the prediction accuracy by feeding more data and
upgrading the RC car hardware platform to enable more precise steering
angle control.

>>>>

%% As future work, we plan to improve the prediction accuracy of the
%% system by feeding more data in training.
As future work, we plan to investigate ways to reduce computational
and memory overhead of DNN inferencing and to evaluate the
effectiveness of FPGA and other specialized accelerators.

%% we plan to apply shared DRAM resource management
%% techniques~\cite{Yun2013,yun2014rtas} on the DeepPicar platform and
%%evaluate their impact on the real-time performance of the system.

%% and
%% upgrading the RC car hardware platform to enable more precise steering
%% angle control
